---
title: "American National Election Survey (ANES) data"
author: "Nathan Alexander"
output: html_document
---

There are at least two packages at the time of this file's compilation for the ANES data.

One package `anesr` is located [here](https://github.com/jamesmartherus/anesr).

You can find the documentation for the other package `anesrake` [here](https://cran.r-project.org/web/packages/anesrake/anesrake.pdf).

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(devtools)

install_github("jamesmartherus/anesr", repos = "http://cran.us.r-project.org")
install.packages("anesrake", repos = "http://cran.us.r-project.org")
```

A few examples on how to access data and use the package are provided below:

## ANESR

```{r, eval=FALSE}
library(anesr)
data(package="anesr") #View a list of available datasets

data(timeseries_cum) #Load Time Series Cumulative File (1948-2016)
data(timeseries_cum_doc) #Load documentation for the Time Series Cumulative File

data(pilot_2018) #Load 2018 Pilot Study
```

View the documentation for the data. Be sure to put `echo=F` to remove the very long output.

```{r, echo=F, eval=FALSE}
timeseries_cum
timeseries_cum_doc
```

View the pilot data from 2018.

```{r, eval=FALSE}
df <- pilot_2018
df
```

## ANESRAKE

```{r, eval=FALSE}
library(anesrake)
```
