---
title: "General Social Survey (GSS) data"
author: "Nathan Alexander"
output: html_document
---

The package `gssr` is located [here](https://kjhealy.github.io/gssr/).

We will run through a sample analysis using the `gssr` package.

```{r}
#| echo: true
#| output: false
#| warning: false
# install packages
install.packages("remotes", repos = "http://cran.us.r-project.org")
install.packages("tidyverse", repos = "http://cran.us.r-project.org")
install.packages("tidyr", repos = "http://cran.us.r-project.org")

# Install 'gssr' from 'ropensci' universe
install.packages('gssr', repos =
  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

# Also recommended: install 'gssrdoc' as well
install.packages('gssrdoc', repos =
  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

# load libraries
library(gssr)
library(dplyr)
library(tidyr)
```

---

#### Explore variables in the GSSR documentation file

```{r}
#| echo: true
#| output: true
#| warning: false
# load the master documentation files
data(gss_all) # note that this is a large file of all GSS data
data(gssrdoc) # this is the documentation for the GSS data
```

---

#### Examine the years available for any given variable

```{r}
# check which years your variables are available
gss_which_years(gss_all, race)
gss_which_years(gss_all, sex)
gss_which_years(gss_all, hrs2)
gss_which_years(gss_all, educ)

# when you want to return information for multiple variables
gss_all %>%
  gss_which_years(c(race, sex, hrs2, educ)) %>%
  print(n = Inf)
```

---

```{r}
#| echo: true
#| output: true
#| warning: false
# loading single year data
gss12 <- gss_get_yr(2012)
gss22 <- gss_get_yr(2022)

# View(gss22)
```

---

```{r}
#| echo: true
#| output: true
#| warning: false
# select your variables for a single year using cross sectional file
gss12 %>% 
  select(race, sex, hrs2, educ, wtssall)

gss12 %>% 
  select(race, sex, hrs2, educ, wtssall) %>% 
  count(race, sex, wtssall)
```

---

# Get data via `gss_all` data set

We can use a more direct method but we must be careful about our outputs.

```{r}
#| echo: true
#| output: true
#| warning: false
gss_all %>% 
  select(year, race, sex, hrs2, educ, wtssall, wtss, wtssnr) %>% 
  filter(year == 2014) %>% 
  drop_na() -> df

sapply(df, function(x) sum(is.na(x))) # take note that the hrs2 variable will cause issue
df
```

---

# Model 1

---

```{r}
#| echo: true
#| output: true
#| warning: false
# select your variables for a single year using gss_all
gss_all %>% 
  filter(year == 2012) %>% 
  select(race, sex, hrs2, educ, wtssall, wtss, wtssnr)

gss_all %>% 
  filter(year == 2018) %>% 
  select(race, sex, hrs2, educ, wtssall, wtss, wtssnr)
```


---

#### Cleaning data

```{r}
#| echo: true
#| output: true
#| warning: false
# clean data for analysis
gss12 %>% 
  select(race, sex, hrs2, educ, wtssall, wtss, wtssnr) %>% 
  drop_na() -> gss12_analysis
```

---

```{r}
#| echo: true
#| output: true
#| warning: false
# check for missing values
sapply(gss12_analysis, function(x) sum(is.na(x)))
str(gss12_analysis)
```


---

```{r}
#| echo: true
#| output: true
#| warning: false
# exploratory analysis
gss12 %>% 
  select(race, sex, hrs2, educ, wtssall, wtss, wtssnr) %>% 
  count(race, sex)

gss12_analysis %>% 
  count(race, sex)
```


---

```{r}
#| echo: true
#| output: true
#| warning: false
model <- lm(gss12_analysis$hrs2 ~ gss12_analysis$educ, weight = gss12_analysis$wtssall)
summary(model)
```


---

# Plot the variable relationships 

Here, I use some dated code, namely the base R abline() function discussed in class.

```{r}
#| echo: true
#| output: true
#| warning: false
# Plot the regression line
plot(gss12_analysis$educ, gss12_analysis$hrs2)

# Some alternate code
plot(gss12_analysis$educ, gss12_analysis$hrs2, 
     main="Education vs Hours Worked",
     xlab="Years of Education", 
     ylab="Hours Worked")

# Add the regression line
abline(lm(hrs2 ~ educ, data = gss12_analysis), col = "red")
```

---

Here, I use some more frequent code with additional parameters given that we've made a ggplot object. The code also provides some labels to our plot.

# Plot the variable relationships 
```{r}
#| echo: true
#| output: true
#| warning: false

# Create a basic scatter plot
plot(gss12_analysis$educ, gss12_analysis$hrs2, 
     main="Education vs Hours Worked",
     xlab="Years of Education", 
     ylab="Hours Worked",
     pch=16,  # Use filled circles for points
     col="blue",  # Make the points blue
     cex=0.8)  # Slightly smaller point size

# Add the regression line
abline(model, col = "red", lwd = 2)

# Add a legend
legend("topleft", legend=c("Data points", "Regression line"),
       col=c("blue", "red"), pch=c(16, NA), lty=c(NA, 1))

# Add text for R-squared value
r_squared <- summary(model)$r.squared
text(x = max(gss12_analysis$educ), y = min(gss12_analysis$hrs2), 
     labels = paste("R-squared =", round(r_squared, 3)), 
     pos = 2)
```
---

# Examine residuals

```{r}
#| echo: true
#| output: true
#| warning: false
# Check residuals
resids <- residuals(model)
plot(gss12_analysis$hrs2, resids)

```

---

```{r}
#| echo: true
#| output: true
#| warning: false
# Diagnostic plots
par(mfrow=c(2,2))
plot(model)
```

---

```{r}
#| echo: true
#| output: true
#| warning: false
# Check significance of predictor
anova(model)
```

---

# Model 2

```{r}
# run your model and diagnostic plots
model2 <- lm(df$educ ~ df$hrs2)
summary(model2)

pred <- predict(model2) 

# Plot the scatterplot
plot(df$educ, df$hrs2)

# Check residuals
resids2 <- residuals(model2)
plot(df$educ, resids2)

# Diagnostic plots
par(mfrow=c(2,2))
plot(model2)

# Check significance of predictor
anova(model2)
```

---

# Preview of multivariate analysis

Try to examine the assumptions of the multivariate analysis and model below.

```{r}
## conducting a multivariate linear analysis
mlm1 <- lm(cbind(hrs2, educ) ~ race + sex, data = df, weight = wtssall)
summary(mlm1)
```

